---
title: "Data 607 Assignment 2 Attempt"
author: "Keith DeNivo"
date: "2024-02-04"
output: html_document
---
http://rpubs.com/divide_by_zero/1145333

<header>Intro.<header>
After using MySQLconnection to generate and post the movie ratings table into cunydata607sql.mysql.database.azure.com an attempt was made to connect to cunydata607sql.mysql.database.azure.com using R/Rstudio. Ideally the table would then be imported and converted to an R dataframe.
```{r packages}
library(RODBC)
library(RSQLite)
library(RMySQL)
library(DBI)
library(odbc)
library(dplyr)
library(dbplyr)
library(RPostgres)
library(sqldf)
```
<header>Packages:
A variety of packages was used to attempt to connect to cunydata607sql.mysql.database.azure.com.

``` {r connectionAttempts}


#odbcDriverConnect(connection = "Driver=SQL #Server;Server=cunydata607sql.mysql.database.azure.com;Port=3306;Uid=keith.denivo49#;Pwd=keith.denivo49;Encrypt=yes;TrustServerCertificate=no;Connection Timeout=30;")  
#ODBC connection failed

#conn <- DBI::dbConnect(
 #   odbc::odbc(),
 #   Driver = "SQL Server",
 #   Server = "cunydata607sql.mysql.database.azure.com",
   
#    uid = "keith.denivo49",
 #   pwd = "keith.denivo49",
#    
#  )
#connection failed



mysqlconnection = dbConnect(RMySQL::MySQL(),
                            server = "cunydata607sql.mysql.database.azure.com",
                            host="cunydata607sql.mysql.database.azure.com",
                            port=3306,
                            user="keith.denivo49",
                            password="keith.denivo49")
#runs, but was unable to access the table.

 DBIconnect <- DBI::dbConnect(RSQLite::SQLite(), "cunydata607sql.mysql.database.azure.com")
#runs but was unsure if it accomplishes anything
 
liteconn <- dbConnect(RSQLite::SQLite(), "cunydata607sql.mysql.database.azure.com")
#runs


#DBI_Connection <- DBI::dbConnect(odbc(), 
#                      driver = "SQL Server",
#                      server="cunydata607sql.mysql.database.azure.com",
#                      Uid="keith.denivo49",
 #                           Pwd="keith.denivo49")
#connection failed

#dt1 <- sqlFetch(channel=mysqlconnection, sqtable = "movie_rating") 
# an attempt to fetch the table.

mysqlconnection <- dbConnect(RMySQL::MySQL(),
                           # server = "cunydata607sql.mysql.database.azure.com",
                            host="cunydata607sql.mysql.database.azure.com",
                            port=3306,
                            user="keith.denivo49",
                            password="keith.denivo49")
#runs 

mysqliteconnection <- dbConnect(SQLite(),
                            server = "cunydata607sql.mysql.database.azure.com",
                            #host="cunydata607sql.mysql.database.azure.com",
                            port=3306,
                            user="keith.denivo49",
                            password="keith.denivo49")
#runs without errors

conn = dbConnect(SQLite(), server="cunydata607sql.mysql.database.azure.com",user='keith.denivo49',port=3306, password='keith.denivo49')
#runs without errors

#conn = dbConnect(RMySQL::MySQL(), #server="cunydata607sql.mysql.database.azure.com",user='keith.denivo49',port=3306, #password='keith.denivo49')
#another attempt to connect to the server

unique(odbcListDrivers()[[1]])
#check if the correct driver is SQL Server

#odbcconn <- odbc::dbConnect(odbc(), 
 #                           Driver = "SQL Server", 
 #                           Server = "cunydata607sql.mysql.database.azure.com",
 #                           Uid="keith.denivo49",
 #                           Pwd="keith.denivo49")
#errors out

#rscon <- DBI::dbConnect(odbc::odbc(),
 #                       Driver    = "SQL Server", 
  #                      Server    = "cunydata607sql.mysql.database.azure.com",
 #                       Database  = "My Database",
 #                       UID       = "keith.denivo49",
 #                       PWD       = "keith.denivo49"
 #                       #,Port      = 3306,
 #                       #Port = 1433
 #                       )
#Error: nanodbc/nanodbc.cpp:1138: 01000
#[Microsoft][ODBC SQL Server Driver][DBNETLIB]SQL Server does not exist or access denied. 
#[Microsoft][ODBC SQL Server Driver][DBNETLIB]ConnectionOpen (Connect())

#dbListTables(conn)
# dbListTables(mysqlconnection) #could not run statement: No database selected

dbListTables(mysqliteconnection)
#runs but only lists character (0) for a table

#dbReadTable(conn, movie_rating)
#dbReadTable(mysqliteconnection)
#attempts to read in the table using the connections that did not error out

#data <- dbGetQuery(conn , "SELECT * FROM movie_rating")
#data <- dbGetQuery(conn , "SELECT * FROM movie_rating")
#data <- dbGetQuery(mysqliteconnection , "SELECT * FROM #`keith.denivo49`.movie_rating")
#attempts to read in the table using the connections that did not error out

src_dbi(conn)

src_dbi(liteconn)
src_dbi(mysqliteconnection)
dbListTables(mysqliteconnection)
#shows additional information about the connection to the database

#src_dbi(mysqlconnection) #errors out
#dbListFields(conn)

#RMySQL::dbListTables(mysqlconnection, sqldf('SELECT * FROM movie_rating;'))
#RMySQL::dbListTables(mysqliteconnection)
#errors out


#con <- dbConnect(RPostgres::Postgres(),dbname = 'keith.denivo49', 
#                 host = 'cunydata607sql.mysql.database.azure.com', # 
#                 port = 5432, # or any other port specified by your DBA
#                user = 'keith.denivo49',
 #                password = 'keith.denivo49')
#Error: connection to server at "cunydata607sql.mysql.database.azure.com" #(20.84.210.207), port 5432 failed: Connection timed out (0x0000274C/10060)
#	Is the server running on that host and accepting TCP/IP connections?

```
<header>information:
Any attempt to connect to cunydata607sql.mysql.database.azure.com resulted in failure. Or at least a failure to recognize the table was was supplied by mySQLworkbench scripts.

#```{sql }

#USE cunydata607sql.mysql.database.azure.com;
#knitr::opts_chunk$set(connection = mysqliteconnection);

#SELECT * FROM movie_rating;
#SELECT * FROM `keith.denivo49`.movie_rating";

#attempts to connect to database using sql script language

#```


<header>Conclusion:

Initially when the table was constructed in mySQL workbench "null" was used in the table for the missing movie ratings in the hopes that R would be able to identify that there was data missing and to not count them for operations like mean. 
