---
title: "data 607 assignment 3"
author: "Keith DeNivo"
date: "2024-02-10"
output: html_document
---

https://rpubs.com/divide_by_zero/1147572
Note on question 4 html rpubs does not show full regex expressions, it deletes the "*" 


```{r library}

library(plyr)
library(tidyverse)
library(dplyr)
library(ggpubr)
library (readr)
library(RCurl)
library(rvest)

```
 
 <b><header>Introduction:</header></b>
 Loaded what was believed to be relevant libraries. <br>
 Homework questions and answers: <br>
 Question 1 <br>
 Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r read data}


c <- getURL("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")
major_list_df <- data.frame(read.csv(text = c) )
#put the data into a dataframe
major_list <- as.list(major_list_df[,2])
glimpse(major_list)
#puts the majors column into a list of majors
```
<header>Question 1 continued</header>
Data was scraped from the web into a dataframe.  the majors in the dataframe were then converted into a list.

"provide code that identifies the majors that contain either "DATA" or "STATISTICS""
```{r Q1}


data_maj <- grep(pattern = 'data', major_list, value = TRUE, ignore.case = TRUE)
# picks the majors with "data" in its string.
stats_maj <- grep(pattern = 'statistics', major_list, value = TRUE, ignore.case = TRUE)
#picks the majors with "statistics" in its string
data_or_stat <- c(data_maj, stats_maj)
#combines the statistics majors and the data majors into one list
print(data_or_stat)

```
<header>Question1 continued</header>
The statistics majors were pulled from the list, then the data majors were pulled.  The lists were then combined into one list.


<header>Question2:</header>
Write code that transforms the data below<br>
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")
```{r fruits Q2}

answer <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")
print(answer)
#the answer to compare results to

fruity <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange" [5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"   [9] "elderberry"   "lime"         "lychee"       "mulberry"     [13] "olive"        "salal berry"'
#assuming the "data" was a single string
print(fruity)
fruity1 <- str_replace_all(fruity, "[:digit:]", "")
fruity2 <- gsub('[[]]', '', fruity1)
#removing all numbers and brackets
fruity_comma <- gsub('" ', '",', fruity2)
#add commas in between quotes
fruity_comma1 <- gsub('"', '', fruity_comma)
#remove the quotes
print(fruity_comma1)
#check quotes were removed
fruity_list <- as.list(str_split(fruity_comma1,",")[[1]])
#split the string so it fits into a list
fruity_list2 <- trimws(fruity_list)
#remove the blank spaces in the beginning and end of strings
print(fruity_list2)
#check list

attempt1 <- fruity_list2
#does the list match the answer?
attempt1 == answer
#list successfully matches answer

#assuming the data was in a dataframe
fruity1 <- data.frame("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")  



fruitys <- c(t(fruity1))
#Converting dataframe into vector/list.
#read the dataframe into a list row by row
attempt2 <- fruitys
#renaming list to attempt2
print(attempt2)

attempt2 == answer
#list successfully matches answer
```
<header>Question 3:</header>
Describe, in words, what these expressions will match:
(.)\1\1 
<br>
this would be rewritten as (.)(.)(.) where (.)=(.)=(.) it will check for any character as denoted by "." it will return parts of a string with three repeating characters "aaa", "111" etc. 
<br>
"(.)(.)\\2\\1"
<br>
returns parts of a string that has this pattern  (1char)(2char)(2char)(1char) like "abba"
<br>
(..)\1
<br>
returns any two characters that repeat together (char1char2)(char1char2) like "toto" or "1111"
<br>
"(.).\\1.\\1"
<br>
returns three of the same character with any character in between each of the same characters or char1 anycharacter char1 anycharacter char1 like "c1chc"
<br>
"(.)(.)(.).*\\3\\2\\1"
<br>
returns any three characters followed by any length string of characters followed by the same initial three characters in reverse order char1 char2 char3 anycharacters char3 char2 char1 such as "123abcde321"

```{r testing regex}

randomstr <- c("aaa", "a111b","fjdkfjd","ckkc","ckck","abeethcba", "abbae1221", "tttt1t1t1", "b3b6b567", "abb222bba222", "abc123456cba")
#strings to test the regex expressions on
x <-  str_view(randomstr, '(.)\\1\\1')
print('(.)\\1\\1')
print(x)
#examples of the first expression
x1 <- str_view(randomstr, "(.)(.)\\2\\1") 
print("(.)(.)\\2\\1")
print(x1)
#examples of the second expression
x2 <- str_view(randomstr, "(..)\\1") 
print("(..)\\1")
print(x2)
#examples of the third expression
x3 <- str_view(randomstr, "(.).\\1.\\1") 
print("(.).\\1.\\1")
print(x3)
#examples of the fourth expression
x4 <- str_view(randomstr, "(.)(.)(.).*\\3\\2\\1") 
print("(.)(.)(.).*\\3\\2\\1")
print(x4)
#examples of the fifth expression

```
<header>Question 4:</header>
 Construct regular expressions to match words that:
<br>
Start and end with the same character.
<br>
"^(.).*//1$" returns strings that start and end with the same character.
<br>

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
<br>
"(.*)(\\w)(\\w).*\\2\\3.*" returns a string of any length that contains repeating pair of letters
<br>

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
<br>
"(.*)(\\w).*\\2.*\\2.*" Returns a string of any length that contains three repeated characters anywhere in the word.
<br>

```{r testing regex expressions Q4}

randomstr <- c("aaa", "a111b","fjdkfjd","ckkc","ckck","abeethcba", "abbae1221", "tttt1t1t1", "b3b6b567", "abb222bba222", "abc123456cba", "church","eleven","beeen")
#strings to test the regex expressions
x5 <-  str_view(randomstr, "^(.).*\\1$")
print("^(.).*\\1$")
print(x5)
x6 <- str_view(randomstr, "(.*)(\\w)(\\w).*\\2\\3.*") 
print("(.*)(\\w)(\\w).*\\2\\3.*")
print(x6)

x7 <- str_view(randomstr, "(.*)(\\w).*\\2.*\\2.*") 
print("(.*)(\\w).*\\2.*\\2.*")
print(x7)


```
<b><header>Conclusion:</header></b>
Answered questions pertaining to selecting segments from strings.
