---
title: "Data 607 Week 5 Assignment"
author: "Keith DeNivo"
date: "2024-02-24"
output: html_document
---
<b><header>Introduction:</header></b><br>
https://rpubs.com/divide_by_zero/1153055
A CSV was created to match the picture in the assignment using excel.  The file was then put into github.  The csv was scraped using R and put into a data frame.  The data frame was put into a tidy format.  Some basic information on the delays between the two airlines was conducted.
```{r Library}
library(dplyr)
library(RCurl)
library(tidyr)
library(tidyverse)
library(reshape2)
```
Libraries Loaded
``` {r csv scrape}
c <- getURL("https://raw.githubusercontent.com/division-zero/Data607/main/Week%205%20Assignment/airline.csv")
#read in the raw file
airline_info_df <- data.frame(read.csv(text = c ))
#put the csv into a dataframe
head(airline_info_df)
```
CSV from github was scraped.

```{r rename and format}

airline_info_df[2,1] <- "ALASKA"
airline_info_df[5,1] <- "AM WEST"
#add airlines to their corresponding rows
airline_info_df <- airline_info_df[-c(3), ]
#remove blank row
head(airline_info_df)

```
<b><header>Fixing original dataframe:</header></b><br>
The missing information was added to the dataframe.  There were two rows that did not mention which airline the data was for.  The blank row was deleted.
```{r arrange}
airline_info_df <- airline_info_df |> rename(Airline = 1, Time = 2)
# rename the columns to something useful
airline_info_df <- airline_info_df |> pivot_longer(Los.Angeles:Seattle, values_to = "frequency", values_drop_na = TRUE)
#pivot the data to a long format by destination.  The number of flights were put into "frequency"
airline_info_df <- airline_info_df |> rename(Destination = 3)
#column was renamed to destination
head(airline_info_df)

```
<b><header>Pivot by destination:</header></b><br>
Data was converted into a long format using pivot by destination.
```{r rearrange}

head(airline_info_df)

airline_info_df <- pivot_wider(airline_info_df, 
                          names_from = Time, 
                          values_from = frequency)
# in order to make the data more in line with a tidy format the on time and delay flights were converted into columns
head(airline_info_df)
```
<b><header>Pivot for each observation:</header></b><br>
Data was pivoted again to include the on time flight and delay flights numbers for each airline and destination. This is the data frame that was written as a csv at the end.

```{r analysis}


airline_info_df_analysis <- mutate(airline_info_df, '%_delays' = airline_info_df$delayed / (airline_info_df$`on time`+airline_info_df$delayed) *100 )
airline_info_df_analysis %>% tbl_df %>% print(n=Inf)
#what percent of flights are delayed for each destination?
alaska_only_df <- airline_info_df_analysis |> 
  filter(Airline == 'ALASKA') 
#df with only Alaska airline in it
mean(alaska_only_df$`%_delays`)
#what is the average percent delay for Alaska?
am_west_only_df <- airline_info_df_analysis |> 
  filter(Airline == 'AM WEST')
#df with only AM WEST info
mean(am_west_only_df$`%_delays`)
#what is the average percent delay for AM WEST?
AM_W_On_time_total <- sum(am_west_only_df$`on time`)
AM_W_delayed_total <- sum(am_west_only_df$delayed)
ALASKA_On_time_total <- sum(alaska_only_df$`on time`)
ALASKA_delayed_total <- sum(alaska_only_df$delayed)
AM_W_delay_percent <- AM_W_delayed_total/(AM_W_delayed_total+AM_W_On_time_total)*100
ALASKA_delay_percent <- ALASKA_delayed_total/(ALASKA_delayed_total+ALASKA_On_time_total)*100
#computing the average percent delay adding all flight together for each airline.
print('Average delay % across all flights for ALASKA')
print(ALASKA_delay_percent)
print('Average delay % across all flights for AM WEST')
print(AM_W_delay_percent)
ggplot( data = airline_info_df_analysis, aes(x = Airline,y = `%_delays`)) + geom_boxplot()  
write.csv(airline_info_df, 'tidy_airline.csv')
```
<b><header>Conclusion:</header></b><br>
Alaska has smaller percentage of delays to each destination.  AM West has a smaller percentage of delays overall for all flights ignoring destination.  This discrepancy is explained by the fact that AM west has the most flights to Phoenix, which has a low % delay. airline_info_df_analysis shows the percent delay for each destination for each airline.